---
title: "`ipeds`: R Package for Accessing the Integrated Postsecondary Education Data System from R"
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- 
[![Build Status](https://api.travis-ci.org/jbryer/ipeds.svg)](https://travis-ci.org/jbryer/ipeds?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/ipeds)](http://cran.r-project.org/package=ipeds)
![Downloads](http://cranlogs.r-pkg.org/badges/ipeds)
-->

#### Installation

<!--
The `ipeds` package can be downloaded from [CRAN](https://cran.r-project.org) as follows:

```{r, eval = FALSE}
install.packages('ipeds')
```
-->

To install the latest development version from Github use the `remotes`` package:

```{r, eval = FALSE}
remotes::install_github('jbryer/ipeds')
```

This package requires [mdbtools](https://github.com/brianb/mdbtools). The following commands will install `mdbtools` on Mac:

```
ruby -e "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/master/install)" < /dev/null 2> /dev/null \n')
brew install mdbtools
```

#### Using the `ipeds` package

The vignette also has useful information for getting started:

```{r, eval = FALSE}
vignette('ipeds')
```

The `available_ipeds` will return a `data.frame` indicating which databases are available for download and my have already been downloaded.

```{r}
ipeds::available_ipeds()
```

The `download_ipeds` will download an IPEDS database for a given year.

```{r}
ipeds::download_ipeds(2021)
```

The `load_ipeds` will return a list of all the survey tables (as `data.frame`s) for the given year.

```{r}
ipeds2021 <- ipeds::load_ipeds(2021)
names(ipeds2021)
```

The `ipeds_help` function will return the data dictionary for the given year.

```{r, eval=FALSE}
View(ipeds_help(2021))
```

If the `table` parameter is specified, then the data dictionary for the given survey is returned (i.e. the variables in that table, see `data(surveys)` for the available survey IDs).

```{r, eval=FALSE}
View(ipeds_help(table = 'HD', year = 2021))
```

To load a specific table, the `ipeds_survey` function will return a `data.frame` with the data.

```{r}
hd2021 <- ipeds::ipeds_survey(table = 'HD', year = 2021)
names(hd2021)
```

Mapping variable factors.

```{r}
names(hd2021) <- tolower(names(hd2021))
hd2021 <- ipeds::recodeDirectory(hd2021)

IvyLeague <- c("186131","190150","166027","130794","215062","182670","217156","190415")
hd2021.ivy <- hd2021[which(hd2021$unitid %in%IvyLeague),]
p <- hd2021.ivy[, c("instnm", "webaddr", "stabbr", "control")]
names(p) <- c("Institution", "Web Address", "State", "Sector")
p
```

Use the enrollment survey data.

```{r}
enrollment <- ipeds::ipeds_survey(table = 'EFFY', year = 2021)
names(enrollment) <- tolower(names(enrollment))

enrollment <- enrollment[which(enrollment$unitid %in% c(IvyLeague) ),]
enrollment <- merge(enrollment, hd2021[, c("unitid", "instnm", "control")], by = "unitid", all.x = TRUE, sort = FALSE)
enrollment <- enrollment[which(enrollment[,"effylev"] == 1),] # Level 1 is Undergraduate

p <- enrollment[, c("instnm", "efytotlt", "control")]
names(p) <- c("Institution", "Total Undergraduate Enrollment", "Sector")
p
```


